---
title: "16S Bacterial Community Analysis 2026"
author: "Dan Shelley"
date: "`r Sys.Date()`"
output: html_document
---

################################################################################
#                                                                              #
#            **IT IS RECOMMENDED TO RUN THIS RMD CHUNK BY CHUNK,**             #
#                                                                              #
#         **THIS MARKDOWN CAN ALSO FUNCTION AS A CLASSIC 16S ANALYSIS**        #
#                                                                              #
################################################################################

################################################################################
#                                                                              #
#                       **Methodology & Package Summary**                      #
#                                                                              #
#        * Phyloseq is used to generate a Phyloseq object from Qiime2          #
#          outputs.                                                            #
#                                                                              #
#        * Taxa filtering: No Filtering for Taxa                               #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
################################################################################

################################################################################
#                                                                              #
#                            **Loading Libraries**                             #
#                                                                              #
################################################################################

```{r library.loading, include=FALSE}
#devtools::install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
devtools::install_github("gauravsk/ranacapa")

library(dplyr)          #v1.2.0   - Used for data cleaning and manipulation.
library(randomForest)   #v4.7-1.2 - Random Forest Model (Not Used in 16S Script).
library(RColorBrewer)   #v1.1-3   - 
library(vegan)          #v2.7-3   - Plotting Random Forest MDS.
library(tidyr)          #v1.3.2   - 
library(ggplot2)        #v4.0.2   - Generic Plotting.
library(Polychrome)     #v        - Colour palette for taxa bar plot.
library(ggpubr)         #v        - 
library(FSA)            #v        - 
library(stringr)        #v        - 
library(ggh4x)          #v        -
library(phyloseq)       #v1.50.0  - Generating objects to store Qiime2 outputs.
library(microbial)      #v        - Phyloseq disperision statistical test.
library(pairwiseAdonis) #v        - Pairwise PERMANOVA functions.
```

################################################################################
#                                                                              #
#                          **Defining Input Files**                            #
#                                                                              #
################################################################################

```{r}
ref_metadata_path        <- "C:/Users/Danie/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/Add to Github/DCWW_HRUU_Qiime2_Output/metadata.csv"
ref_otu_table_path       <- "C:/Users/Danie/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/Add to Github/DCWW_HRUU_Qiime2_Output/feature-table.tsv"
ref_otu_fasta_path       <- "C:/Users/Danie/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/Add to Github/DCWW_HRUU_Qiime2_Output/aligned-dna-sequences.fasta"
ref_taxonomy_table_path  <- "C:/Users/Danie/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/Add to Github/DCWW_HRUU_Qiime2_Output/taxonomy.tsv"
ref_tree_path            <- "C:/Users/Danie/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/Add to Github/DCWW_HRUU_Qiime2_Output/tree.nwk"

output_directory <- "C:/Users/Danie/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/MICROBIAL COMMUNITY ANALYSIS/output/"
```

################################################################################
#                                                                              #
#                       **Generating Phyloseq Object**                         #
#                                                                              #
################################################################################

```{r read_in_reference_data}
feature_table <- read.table(
  ref_otu_table_path,
  header = TRUE,
  row.names = 1,
  sep = "\t"
)

feature_table <- as.matrix(feature_table)
feature_table <- otu_table(feature_table, taxa_are_rows = TRUE)


taxonomy_table <- read.table(
  ref_taxonomy_table_path,
  sep = "\t",
  header = TRUE,
  row.names = 1
)

taxonomy_split <- strsplit(as.character(taxonomy_table$Taxon), ";")

taxonomy_split_formatted <- lapply(taxonomy_split, function(x) {
  length(x) <- 7
  return(x)
})

taxonomy_mtx <- as.matrix(do.call(rbind, taxonomy_split_formatted))

# Remove prefixes like "d__", "p__" etc.
taxonomy_mtx <- sub("^[a-z]__*", "", taxonomy_mtx)  # removes d__, p__, etc.
taxonomy_mtx[taxonomy_mtx == ""] <- NA # now we fill in the remaining empty gaps with NAs so R doesn't get confused

# Set column names (use "Domain" or "Kingdom" as you prefer)
colnames(taxonomy_mtx) <- c("Domain",
                            "Phylum",
                            "Class",
                            "Order",
                            "Family",
                            "Genus",
                            "Species")
# Add back the ASV IDs which were lost from our original data frame
rownames(taxonomy_mtx) <- rownames(taxonomy_table)

# Convert to phyloseq tax_table object
taxonomy_table <- phyloseq::tax_table(taxonomy_mtx)

metadata_table <- read.csv(
  ref_metadata_path,
  row.names = 1,
  na.strings = "",
  stringsAsFactors = TRUE
)

# Match Metadata Sample IDs to those in the feature table
metadata_table <- metadata_table[rownames(metadata_table) %in% colnames(feature_table), ]
metadata_table$Sample.name <- rownames(metadata_table)

metadata_table <- sample_data(metadata_table)

rooted_tree <- read_tree(ref_tree_path)

psq_obj_ref <- phyloseq(feature_table, taxonomy_table, metadata_table, rooted_tree)
```

################################################################################
#                                                                              #
#                              **Data Cleaning**                               #
#                                                                              #
################################################################################

```{r check_ref_psq}

sample_data(psq_obj_ref)$anon_location <- recode(
  sample_data(psq_obj_ref)$Location,
  "Alaw" = "WTW1",
  "Hirwaun" = "WTW2",
  "Maerdy" = "WTW3",
  "Pontsticill" = "WTW4",
  "Tynywaun" = "WTW5",
  "Cefni" = "WTW6",
  "Glascoed" = "WTW7",
  "Hurleston" = "Hurleston",
  "Sluvad" = "WTW8",
  "Preseli" = "WTW9"
)

sample_data(psq_obj_ref)$anon_location <- as.character(sample_data(psq_obj_ref)$anon_location)
sample_data(psq_obj_ref)$anon_location <- factor(
  sample_data(psq_obj_ref)$anon_location,
  levels = c(
    "WTW1",
    "WTW2",
    "WTW3",
    "WTW4",
    "WTW5",
    "WTW6",
    "WTW7",
    "WTW8",
    "WTW9",
    "Hurleston"
  )
)

sample_data(psq_obj_ref)$Sample.details <- recode(
  sample_data(psq_obj_ref)$Sample.details,
  "Clarified.Water" = "CLR-Water",
  "Clarified-Water" = "CLR-Water",
  "clarifier-sludge" = "CLR-Sludge",
  "Clarifier-sludge" = "CLR-Sludge",
  "COCODAF" = "COCODAF",
  "COCODAF-sludge" = "COCODAF-Sludge",
  "COCODAF-backwash" = "COCODAF-Backwash",
  "DAF-sludge" = "DAF-Sludge",
  "DAF" = "DAF-Sludge",
  "DAF-blanket" = "DAF-Blanket",
  "DAF.Sludge" = "DAF-Sludge",
  "Filtered-water" = "Filtered-Water",
  "Prefiltered-water" = "Prefiltered-Water",
  "Raw-water" = "RAW-Water",
  "Raw.Water" = "RAW-Water",
  "RGF-backwash" = "RGF-Backwash",
  "RGF-water" = "RGF-Water",
  "RGF.water" = "RGF-Water",
  "Control" = "Control"
)

psq_obj_ref <- subset_samples(psq_obj_ref, !Location %in% c("Negative", "Blank", "Mock-communityA", "Mock-communityB",
                                                            "Sluvad", "Cefni", "Glascoed", "Preseli"))

## Removing taxa which appear a total of 10 or less across all samples
physeq_filtered <- prune_taxa(
  taxa_sums(psq_obj_ref) >= 10,
  psq_obj_ref
)

## Removing Control Groups
yes_vals <- c("Y", "y", "Yes", "yes", "negative", "neg", "true")
sample_ctrl <- "num"

sample_data(psq_obj_ref)$Sequencing.depth <- sample_sums(psq_obj_ref)

psq_obj_ref <- prune_samples(
  sample_data(psq_obj_ref)$Sequencing.depth >= 1000,
  psq_obj_ref
)

```

```{r}

location_palette <- c(
  "WTW1" = "#1b9e07",
  "WTW2" = "#d95f02",
  "WTW3" = "#7570b3",
  "WTW4" = "#e7298a",
  "WTW5" = "brown",
  "WTW6" = "#e6ab02",
  "WTW7" = "#a6761d",
  "WTW8" = "#666666",
  "WTW9" = "#66CCEE",
  "Hurleston" = "#004488"
)

group_palette <- c(
    "RAW-Water"      = "#1b9e77",
    "CLR-Water"      = "#d95f02",
    "DAF-Sludge"     = "#7570b3",
    "RGF"            = "#e7298a",
    "Backwash-Water" = "#e6ab02",
    "Backwash"       = "#1f78b4",
    "Control"        = "#666666"
)
```



################################################################################
#                                                                              #
#       psq_obj_ref              : Full DCWW & Hurleston Dataset, clean        #
#       psq_obj_Hurleston_Clean  : Subset Hurleston Data, clean                #
#                                                                              #
################################################################################


################################################################################


################################################################################
#                                                                              #
#                             **ALPHA DIVERSITY**                              #
#                                                                              #
################################################################################

**(1.) ALPHA RAREFRACTION CURVE OF HURLESTON**

```{r}
# Define depths
lib_sizes <- sample_sums(psq_obj_Hurleston_Clean)
min_lib   <- min(lib_sizes)
max_lib   <- max(lib_sizes)

# Start with your candidate depths
depths <- seq(1, max_lib, by = 1000)
sample_var <- "SampleID"

# Initialize list
richness_list <- list()

for (d in depths) {
  set.seed(2006576)
  rare_dat <- rarefy_even_depth(psq_obj_Hurleston_Clean, sample.size = d, rngseed = 2006576, verbose = FALSE, replace = FALSE)

  richness <- estimate_richness(rare_dat, measures = "Observed")
  richness$SampleID <- rownames(richness)
  richness$Depth <- d
  richness_list[[as.character(d)]] <- richness
}

# Combine results
rare_df <- bind_rows(richness_list)

# Add metadata
metadata <- data.frame(sample_data(psq_obj_Hurleston_Clean))
metadata$SampleID <- rownames(metadata)
rare_df <- left_join(rare_df, metadata, by = "SampleID")

# Summarize mean & SE per group
summary_df <- rare_df %>%
  group_by(Depth, !!sym(sample_var)) %>%
  mutate(
    Richness_mean = mean(Observed, na.rm = TRUE),
    Richness_se   = sd(Observed, na.rm = TRUE) / sqrt(n())
  ) %>%
  ungroup()

Rare.plot.UU <- ggplot(summary_df, aes(x = Depth, y = Richness_mean, group = SampleID, color = Sample.details)) +
  geom_line(linewidth = 1) +
  scale_color_discrete(name = "Sample") +
  scale_x_continuous(breaks = seq(0, 100000, by = 5000)) +
  labs(x = "Sequencing Depth (n)", y = "Mean Observed ASVs (n)") +
  theme_bw()

filename <- paste0(output_directory,"rarefraction_curve_Hurleston.svg")
ggsave(filename, Rare.plot.UU, height = 5, width = 10)
```
**(1.) ALPHA RAREFRACTION CURVE OF DCWW/HURLESTON**

```{r}
# Define depths
lib_sizes <- sample_sums(psq_obj_ref)
min_lib   <- min(lib_sizes)
max_lib   <- max(lib_sizes)

# Start with your candidate depths
depths <- seq(1, max_lib, by = 10000)
sample_var <- "Sample.name"

# Initialize list
richness_list <- list()

for (d in depths) {
  set.seed(2006576)
  rare_dat <- rarefy_even_depth(psq_obj_ref, sample.size = d, rngseed = 2006576, verbose = FALSE, replace = FALSE)

  richness <- estimate_richness(rare_dat, measures = "Observed")
  richness$SampleI.name <- rownames(richness)
  richness$Depth <- d
  richness_list[[as.character(d)]] <- richness
}

# Combine results
rare_df <- bind_rows(richness_list)

# Add metadata
metadata <- data.frame(sample_data(psq_obj_ref))
metadata$SampleID <- rownames(metadata)
rare_df <- left_join(rare_df, metadata, by = "Sample.name")

# Summarize mean & SE per group
summary_df <- rare_df %>%
  group_by(Depth, !!sym(sample_var)) %>%
  mutate(
    Richness_mean = mean(Observed, na.rm = TRUE),
    Richness_se   = sd(Observed, na.rm = TRUE) / sqrt(n())
  ) %>%
  ungroup()

Rare.plot <- ggplot(summary_df, aes(x = Depth, y = Richness_mean, group = Sample.name, color = Sample.details)) +
  geom_line(linewidth = 1) +
  scale_color_discrete(name = "Sample") +
  scale_x_continuous(breaks = seq(0, 500000, by = 5000)) +
  xlim(0,100000) +
  labs(x = "Sequencing Depth (n)", y = "Mean Observed ASVs (n)") +
  theme_bw()

filename <- paste0(output_directory,"rarefraction_curve_Hurleston.svg")
ggsave(filename, Rare.plot, height = 5, width = 10)
```

**Rarifying and subbing out Hurleston + Data Cleaning**
```{r}
psq_obj_ref_rare <- rarefy_even_depth(psq_obj_ref, 12000, rngseed = 202020)
psq_obj_ref_rare_fam <- tax_glom(psq_obj_ref_rare, taxrank = "Family")

## Remove taxa which are only appearing in one sample (experiments were run in triplicate).
psq_obj_Hurleston <- subset_samples(psq_obj_ref_rare_fam, Location == "Hurleston")

## Remove taxa which are only appearing in one sample (experiments were run in triplicate).
psq_obj_Hurleston_Clean <- prune_taxa(taxa_sums(psq_obj_Hurleston) > 0, psq_obj_Hurleston)

sample_data(psq_obj_Hurleston_Clean)$Sample.details <- as.character(sample_data(psq_obj_Hurleston_Clean)$Sample.details)

sample_data(psq_obj_Hurleston_Clean)$Sample.details <- trimws(sample_data(psq_obj_Hurleston_Clean)$Sample.details)

sample_data(psq_obj_Hurleston_Clean)$Sample.details <- recode(
  sample_data(psq_obj_Hurleston_Clean)$Sample.details,
  "RGF1.Backwash" = "Backwash",
  "RGF.Water" = "Backwash-Water"
)

sample_data(psq_obj_Hurleston_Clean)$Sample.details <- as.character(sample_data(psq_obj_Hurleston_Clean)$Sample.details)
sample_data(psq_obj_Hurleston_Clean)$Sample.details <- factor(
  sample_data(psq_obj_Hurleston_Clean)$Sample.details,
  levels = c(
    "RAW-Water",
    "CLR-Water",
    "DAF-Sludge",
    "RGF",
    "Backwash-Water",
    "Backwash",
    "Control"
  )
)
```

################################################################################
#                                                                              #
#           psq_obj_ref_rare : Full Phyloseq Obj with all tax ranks            #
#           psq_obj_ref_rare_fam : Family level Phyloseq Obj                   #
#                                                                              #
################################################################################

**(3.) Calculating Alpha Diversity Distances**

```{r}
group      <- "Sample.details"   # column used for grouping
ps11       <- psq_obj_Hurleston_Clean       # your phyloseq object


# Extract sample metadata
mapping <- phyloseq::sample_data(ps11)

# Make sure all taxa have some non-zero counts
ps11 <- phyloseq::filter_taxa(ps11, function(x) sum(x) > 0, prune = TRUE)

# Rename the grouping column for clarity
colnames(mapping) <- gsub(group, "Comparative_Group", colnames(mapping))

# Create final grouping variable
mapping$Group <- as.factor(mapping$Comparative_Group)

# Convert OTU table into a standard count matrix (samples x taxa)
count <- as.data.frame(t(ranacapa::vegan_otu(ps11)))

# Transpose back for vegan functions that expect taxa as columns
x <- t(count)

# Shannon diversity
Shannon <- vegan::diversity(x, index = "shannon")

# Inverse Simpson
Inv_Simpson <- vegan::diversity(x, index = "invsimpson")

# Species richness (observed)
Richness <- vegan::specnumber(x)

# Prevalence (alternative richness measure)
S2 <- rowSums(x > 0)

# Pielou's evenness
Pielou_evenness <- Shannon / log(Richness)

# Simpson evenness
Simpson <- Inv_Simpson / Richness

# Chao1, ACE, and richness from estimateR
est <- vegan::estimateR(x)
ACE   <- est["S.ACE", ]

# Combine all metrics into a single table
report <- cbind(
  Shannon,
  Inv_Simpson,
  Pielou_evenness,
  Simpson,
  Richness,
  ACE
)

alp <- merge(mapping, report, by = "row.names", all = FALSE)

# Store complete alpha-diversity table
alpha_table <- alp

# Select metrics to test
metrics_to_test <- c(
  "Shannon", "Inv_Simpson", "Pielou_evenness",
  "Simpson", "Richness", "ACE"
)


data_for_test <- cbind(
  data.frame(
    ID = seq_len(nrow(alpha_table)),
    Group = alpha_table$Group,
    SampleID = alpha_table$Row.names,
    Sequencing.depth = alpha_table$Sequencing.depth
  ),
  alpha_table[, metrics_to_test]
)


```

**(4.) Plotting Distance Matrices**

```{r, fig.height=10, fig.width=5}

alpha_metrics <- c(
  "Shannon",
  "Richness",
  "ACE",
  "Simpson",
  "Inv_Simpson",
  "Pielou_evenness"
)


make_alpha_plot <- function(data, metric, palette) {

  # Pairwise Wilcoxon
  sigdf <- pairwise.wilcox.test(
    x = data[[metric]],
    g = data$Group,
    p.adjust.method = "BH"
  )

  pw_df <- as.data.frame(as.table(sigdf$p.value)) %>%
    rename(
      group1 = Var1,
      group2 = Var2,
      p = Freq
    ) %>%
    filter(!is.na(p)) %>%
    mutate(
      p.signif = case_when(
        p <= 0.0001 ~ "****",
        p <= 0.001  ~ "***",
        p <= 0.01   ~ "**",
        p <= 0.05   ~ "*",
        TRUE        ~ "ns"
      )
    ) %>%
    filter(p.signif != "ns")

  # Dynamic y positions
  y_max <- max(data[[metric]], na.rm = TRUE)
  step  <- 0.05 * y_max

  pw_df <- pw_df %>%
    arrange(p) %>%
    mutate(y.position = y_max + row_number() * step)

  # Plot
  ggplot(data, aes(x = Group, y = .data[[metric]], fill = Group)) +
    geom_boxplot() +
    stat_boxplot(geom = "errorbar", width = 0.2) +
    geom_point(aes(color = Group), alpha = 0.8) +
    labs(
      title = metric,
      y = metric,
      fill = "Sample"
    ) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    scale_fill_manual(values = palette) +
    scale_color_manual(values = palette) +
    guides(color = "none") +
    stat_pvalue_manual(
      pw_df,
      label = "p.signif",
      tip.length = 0.01
    )
}


alpha_plots <- lapply(
  alpha_metrics,
  make_alpha_plot,
  data = data_for_test,
  palette = group_palette
)

names(alpha_plots) <- alpha_metrics


final_alpha_plot <- ggarrange(
  plotlist = alpha_plots,
  ncol = 3,
  nrow = 2,
  common.legend = TRUE,
  legend = "right"
)

final_alpha_plot

filename <- paste0(output_directory,"Alpha_Diversity_HRUU.svg")
ggsave(filename, final_alpha_plot, height = 7.5, width = 10)
```


```{r, fig.width=10, fig.height=10}

# Create a list containing distances generated
alpha_list <- c("Shannon",
                "Inv_Simpson",
                "Pielou_evenness",
                "Simpson_evenness",
                "Richness")

# Standard depth for predictions
str.depth <- median(data_for_test$Sequencing.depth)

# Empty plot list to populate
plot_list <- list()

# Loop to predict/standardise distance values by sequencing depth
for (metric in alpha_list) {
  # Build formula dynamically: e.g. Shannon ~ Group + Sequencing.depth
  f <- as.formula(paste(metric, "~ Group + Sequencing.depth"))
  
  # Fit model
  fit <- lm(data = data_for_test, formula = f)
  
  # Name for standardised metric column
  newcol <- paste0("str.", metric)
  
  # Predict standardised values
  data_for_test[[newcol]] <- predict(fit,
                                     newdata = data.frame(Group = data_for_test$Group, Sequencing.depth = str.depth))
  
}

# Function to look all distances to generate alpha diversity plots with statistics
make_alpha_plot <- function(metric, data) {
  
  # Apply "str." to recognise new columns 
  str_metric <- paste0("str.", metric)
  
  # Generate same model as above for emmeans
  m <- lm(as.formula(paste(metric, "~ Group + Sequencing.depth")), data = data)
  
  # Apply pairwise emmeans to the model and compare by 'Group'
  pw_test <- emmeans::contrast(emmeans::emmeans(m, specs = "Group"),
                               method = "pairwise",
                               adjust = "BH")
  
  # Extract key information from the emmeans output to be able to plot
  pw_df <- as.data.frame(pw_test) %>%
    mutate(
      group1 = sub("^(.*) - (.*)$", "\\1", contrast),
      group2 = sub("^(.*) - (.*)$", "\\2", contrast),
      
      # remove parentheses from group names
      group1 = gsub("[()]", "", group1),
      group2 = gsub("[()]", "", group2),
      
      p = p.value,
      p.signif = case_when(
        p <= 0.0001 ~ "****",
        p <= 0.001  ~ "***",
        p <= 0.01   ~ "**",
        p <= 0.05   ~ "*",
        TRUE ~ "ns"
      )
    )
  
  # Filter significance table to remove any non-significance to prevent plotting
  pw_sig <- pw_df %>% filter(p.signif != "ns")
  
  # Scalling feature to apply space for significance bars if any
  base_max <- max(data[[str_metric]], na.rm = TRUE)
  plot_scale <- base_max * 0.15
  
  if (nrow(pw_sig) > 0) {
    pw_sig$y.position <- base_max + seq(plot_scale, by = plot_scale, length.out = nrow(pw_sig))
  }
  
  # ggplot of alpha distances, with expansions for significance bars where needed
  p <- ggplot(data, aes(x = Group, y = .data[[str_metric]])) +
    geom_boxplot(aes(fill = Group),
                 alpha = 0.8,
                 outlier.shape = NA) +
    geom_point(aes(color = Group), size = 3, alpha = 0.8) +
    labs(x = "", y = metric) +
    theme_bw() +
    scale_fill_manual(values = group_palette) +
    scale_color_manual(values = group_palette) +
    scale_y_continuous(expand = expansion(mult = c(0.1, 0.15))) +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5),
      plot.title = element_text(size = 12, face = "bold")
    )
  
  if (nrow(pw_sig) > 0) {
    p <- p + ggpubr::stat_pvalue_manual(pw_sig, label = "p.signif", tip.length = 0.01)
  }
  
  return(p)
}

# Apply the function to the distances within the data
plots <- lapply(alpha_list, make_alpha_plot, data = data_for_test)

# Combine the 5x plots into one figure using ggarrange
combined_plot <- ggpubr::ggarrange(
  plotlist = plots,
  ncol = 3,
  nrow = 2,
  common.legend = TRUE,
  legend = "right"
)

# plot the figure
combined_plot


# ggsave("C:/Users/Dan/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/UnitedUtilities/000 - Reports/Code Repository/Alpha_diversity.svg", 
#        combined_plot, 
#        height = 10, width = 10)

filename <- paste0(output_directory,"alpha_diversity_Hurleston.svg")
ggsave(filename, combined_plot, height = 10, width = 10)

```

################################################################################
#                                                                              #
#                              **BETA  DIVERSITY**                             #
#                                                                              #
################################################################################

**(1.) Hurleston Data**

```{r}
psq_obj_Hurleston_Clean_rel <- normalize(psq_obj_Hurleston_Clean, method = "relative")
psq_obj_Hurleston_Clean_rel <- prune_taxa(taxa_sums(psq_obj_Hurleston_Clean_rel) > 0, psq_obj_Hurleston_Clean_rel)

HRUU_beta <- plotbeta(psq_obj_Hurleston_Clean_rel, group = "Sample.details") + 
  labs(colour = "") 

filename <- paste0(output_directory,"beta_diversity_Hurleston.svg")
ggsave(filename, HRUU_beta, height = 5, width = 7.5)
```

**(2.) PERMANOVA: Statistically comparing microbial community similarities**

```{r}

bact <- data.frame(otu_table(psq_obj_Hurleston_Clean_rel))
metadata <- data.frame(sample_data(psq_obj_Hurleston_Clean_rel))
sample_data(psq_obj_Hurleston_Clean_rel)$Sample.details <- as.character(sample_data(psq_obj_Hurleston_Clean_rel)$Sample.details)
psq_obj_Hurleston_Clean_rel <- subset_samples(psq_obj_Hurleston_Clean_rel, Sample.details !="NA")

## PERMANOVA to determine if there may be a sigdifference between groups and seq depth
dist.bray <- phyloseq::distance(psq_obj_Hurleston_Clean_rel, method = "bray")
ps.adonis <- adonis2(dist.bray ~ Sample.details + Sequencing.depth, data = metadata, perm=9999)

## Statistical approach to compare in-sample dispersion
ps.disper <- betadisper(dist.bray, metadata$Sample.details)
permutest(ps.disper, pair=TRUE)

## Pairwise PERMANOVA to find true differences between groups
pairwise.adonis(dist.bray, factors = sample_data(psq_obj_Hurleston_Clean_rel)$Sample.details)
```

**(3.) Hurleston comparison to DCWW**

```{r}

# (1.) IDENTIFYING TAXA SHARED BETWEEN THE DCWW AND HURLESTON AT THE FAMILY LEVEL
psq_family_RGF <- subset_samples(psq_obj_ref_rare_fam, Sample.details == "RGF")
psq_family_DAF <- subset_samples(psq_obj_ref_rare_fam, Sample.details == "DAF-Sludge")
psq_family_RAW <- subset_samples(psq_obj_ref_rare_fam, Sample.details == "RAW-Water")

# (2.) EXTRACT HURLESTON DATA FROM THE LARGER OBJECT
psq_hurleston <- subset_samples(psq_family_RGF, anon_location == "Hurleston")
psq_hurleston <- prune_taxa(taxa_sums(psq_hurleston) > 0, psq_hurleston)

# (3.) EXTRACT DATA WHICH ISNT HURLESTON
psq_dcww <- subset_samples(psq_family_RGF, anon_location != "Hurleston")
psq_dcww <- prune_taxa(taxa_sums(psq_dcww) > 0, psq_dcww)

# (4.) COMPARE THE TAXA ID AND FINE WHICH ARE SHARED
shared_families <- intersect(
  taxa_names(psq_hurleston),
  taxa_names(psq_dcww)
)

# (5.) COLLAPSE PHYLOSEQ OBJECT BY SHARED TAXA
psq_family_shared <- prune_taxa(shared_families, psq_family_RGF)
psq_family_shared <- prune_taxa(taxa_sums(psq_family_shared) > 0, psq_family_shared)
psq_family_shared

psq_obj_ref_rel <- normalize(psq_family_shared, method = "relative")
```
## Comparing only RGFs at Hurleston

```{r}
psq_hurleston_rel <- normalize(psq_hurleston, method = "relative")
plotbeta(psq_hurleston_rel, group="batch.name") + 
  stat_ellipse(level = 0.95) +
  labs(colour = "RGF ID")

## Full Pairwise: Dispersion is Significant and can comprimise PERMANOVA resutls
ps_tr <- microbiome::transform(psq_hurleston_rel, "clr")
ps_dist_matrix <- phyloseq::distance(psq_hurleston_rel, method ="bray")
vegan::adonis2(ps_dist_matrix ~ phyloseq::sample_data(ps_tr)$batch.name)
pairwise.adonis(ps_dist_matrix, phyloseq::sample_data(ps_tr)$batch.name)

## Pairwise between RGF4 and RGF6 as MH asked
ps_rgf4_6 <- subset_samples(psq_hurleston_rel, batch.name %in% c("RGF6", "RGF4"))
ps_rgf4_6 <- prune_taxa(taxa_sums(ps_rgf4_6) > 0, ps_rgf4_6)

bray_rgf <- distance(ps_rgf4_6, method = "bray")

meta_rgf <- as(sample_data(ps_rgf4_6), "data.frame")

adonis_rgf <- adonis2(
  bray_rgf ~ batch.name,
  data = meta_rgf,
  permutations = 999
)

adonis_rgf

```

```{r}
psq_RGF_rel <- normalize(psq_family_RGF, method = "relative")
plotbeta(psq_RGF_rel, group="anon_location") + 
  stat_ellipse(level = 0.95) +
  labs(colour = "WTW") +
  scale_colour_manual(values = location_palette)

## Full Pairwise: Dispersion is Significant and can comprimise PERMANOVA resutls
ps_tr <- microbiome::transform(psq_RGF_rel, "clr")
ps_dist_matrix <- phyloseq::distance(psq_RGF_rel, method ="bray")
vegan::adonis2(ps_dist_matrix ~ phyloseq::sample_data(ps_tr)$anon_location)
pairwise.adonis(ps_dist_matrix, phyloseq::sample_data(ps_tr)$anon_location)

# Significant Dispersion in RGF Beta Distance: R2 = 0.507, F = 42.151, p = 0.001
# Pairwise Comparisons show all as significant: p = 0.015 (likely due to large dispersion)

```

```{r}
psq_DAF_rel <- normalize(psq_family_DAF, method = "relative")
plotbeta(psq_DAF_rel, group="anon_location") + 
  stat_ellipse(level = 0.95) +
  labs(colour = "WTW") +
  scale_colour_manual(values = location_palette)

## Full Pairwise: Dispersion is Significant and can comprimise PERMANOVA resutls
ps_tr <- microbiome::transform(psq_DAF_rel, "clr")
ps_dist_matrix <- phyloseq::distance(psq_DAF_rel, method ="bray")
vegan::adonis2(ps_dist_matrix ~ phyloseq::sample_data(ps_tr)$anon_location)
pairwise.adonis(ps_dist_matrix, phyloseq::sample_data(ps_tr)$anon_location)
anova(betadisper(ps_dist_matrix, group = phyloseq::sample_data(ps_tr)$anon_location))

# Pairwise Comparisons show significant comparisons: No Significant Comparisons with Hurleston
# Significant Dispersion in DAF samples by locations: F = 4.564, P < 0.01, df = 5
# Significant PERMANOVA results are likely influenced by dispersion
```

```{r}
psq_RAW_rel <- normalize(psq_family_RAW, method = "relative")
plotbeta(psq_RAW_rel, group="anon_location") + 
  stat_ellipse(level = 0.95) +
  labs(colour = "WTW") +
  scale_colour_manual(values = location_palette)

plotbar(psq_RAW_rel, group = "anon_location", level = "Family", top = 20, fontsize.x = 7.5, fontsize.y = 7.5) +
  labs(x = "WTW ID", y = "Relative Abundance")

## Full Pairwise: Dispersion is Significant and can comprimise PERMANOVA resutls
ps_tr <- microbiome::transform(psq_RAW_rel, "clr")
ps_dist_matrix <- phyloseq::distance(psq_RAW_rel, method ="bray")
vegan::adonis2(ps_dist_matrix ~ phyloseq::sample_data(ps_tr)$anon_location)
pairwise.adonis(ps_dist_matrix, phyloseq::sample_data(ps_tr)$anon_location)
anova(betadisper(ps_dist_matrix, group = phyloseq::sample_data(ps_tr)$anon_location))

# Global PERMANOVA is Significant ; R2 = 0.317, F = 2.232, p = 0.001
# Pairwise Comparisons show significant comparisons: No Significant Comparisons with Hurleston
# No Significant Dispersion in Raw water samples by locations: F = 2.467, p = 0.0613, df = 5
# Significant PERMANOVA results are likely true
```


```{r}
bact <- data.frame(otu_table(psq_obj_ref_rel))
metadata <- data.frame(sample_data(psq_obj_ref_rel))
sample_data(psq_obj_ref_rel)$Sample.details <- as.character(sample_data(psq_obj_ref_rel)$Sample.details)
psq_obj_Hurleston_Clean_rel <- subset_samples(psq_obj_ref_rel, Sample.details !="NA")

## PERMANOVA to determine if there may be a sigdifference between groups and seq depth
dist.bray <- phyloseq::distance(psq_obj_Hurleston_Clean_rel, method = "bray")
ps.adonis <- adonis2(dist.bray ~ anon_location, data = metadata, perm=9999)

## Statistical approach to compare in-sample dispersion
ps.disper <- betadisper(dist.bray, metadata$anon_location)
permutest(ps.disper, pair=TRUE)

## Pairwise PERMANOVA to find true differences between groups
pairwise.adonis(dist.bray, factors = sample_data(psq_obj_Hurleston_Clean_rel)$anon_location)
```


**(4.) Statistically compare RGF of different locations**

```{r}
bact <- data.frame(otu_table(psq_obj_ref_rel))
metadata <- data.frame(sample_data(psq_obj_ref_rel))

## PERMANOVA to determine if there may be a sigdifference between groups and seq depth
dist.bray <- phyloseq::distance(psq_obj_ref_rel, method = "bray")
ps.adonis <- adonis2(dist.bray ~ anon_location, data = metadata, perm=9999)

## Statistical approach to compare in-sample dispersion
ps.disper <- betadisper(dist.bray, metadata$anon_location)
permutest(ps.disper, pair=TRUE)

## Pairwise PERMANOVA to find true differences between groups
pairwise.adonis(dist.bray, factors = metadata$anon_location, p.adjust.m = "BH")
```

################################################################################
#                                                                              #
#                           **TAXONOMIC BAR PLOTS**                            #
#                                                                              #
################################################################################

**(1.) Hurleston**

```{r}
Hurleston_bar <- plotbar(psq_obj_Hurleston_Clean_rel, group = "Sample.details", level = "Family", top = 30) + 
  labs(y = "Relative Abundance", x = "Sample Stage", title = "RGF") +
  theme(axis.text.x = element_text(size = 12))

filename <- paste0(output_directory,"bar_plot_Hurleston_RGF.svg")
ggsave(filename, Hurleston_bar, height = 10, width = 10)
```



**(1.) Bar plot of RGF for comparison at different locations**

```{r fig.height=10, fig.width=10}
psq_family_shared_RGF <- prune_taxa(shared_families, psq_family_RGF)
psq_family_shared_RGF <- prune_taxa(taxa_sums(psq_family_shared_RGF) > 0, psq_family_shared_RGF)
psq_family_shared_RGF

psq_family_shared_RGF_rel <- transform_sample_counts(psq_family_shared_RGF, function(x) x / sum(x))


RGF_bar <- plotbar(psq_family_shared_RGF_rel, group = "anon_location", level = "Family", top = 20) + 
  labs(y = "Relative Abundance", x = "Water Treatment Works", title = "RGF") +
  theme(axis.text.x = element_text(size = 12))

filename <- paste0(output_directory,"bar_plot_DCWW_Hurleston_RGF.svg")
ggsave(filename, RGF_bar, height = 10, width = 7.5)
```

```{r}
RGF_beta <- plotbeta(psq_family_shared_RGF_rel, group="anon_location") + 
  stat_ellipse(level = 0.95) +
  labs(colour = "Water Works")+
  scale_colour_manual(values = location_palette)

filename <- paste0(output_directory,"beta_diversity_DCWW_Hurleston_RGF.svg")
ggsave(filename, RGF_beta, height = 5, width = 7.5)
```

**(2.) Bar plot of DAF Sludge for comparison at different locations**

```{r fig.height=10, fig.width=10}
psq_family_shared_DAF <- prune_taxa(shared_families, psq_family_DAF)
psq_family_shared_DAF <- prune_taxa(taxa_sums(psq_family_shared_DAF) > 0, psq_family_shared_DAF)
psq_family_shared_DAF

psq_family_shared_DAF_rel <- transform_sample_counts(psq_family_shared_DAF, function(x) x / sum(x))


DAF_bar <- plotbar(psq_family_shared_DAF_rel, group = "anon_location", level = "Family", top = 20) + 
  labs(y = "Relative Abundance", x = "Water Treatment Works", title = "DAF") +
  theme(axis.text.x = element_text(size = 12))

filename <- paste0(output_directory,"bar_plot_DCWW_Hurleston_DAF.svg")
ggsave(filename, DAF_bar, height = 10, width = 7.5)
```

```{r}
DAF_beta <- plotbeta(psq_family_shared_DAF_rel, group="anon_location") + 
  stat_ellipse(level = 0.95) +
  labs(colour = "Water Works") +
  scale_colour_manual(values = location_palette)

filename <- paste0(output_directory,"beta_diversity_DCWW_Hurleston_DAF.svg")
ggsave(filename, DAF_beta, height = 5, width = 7.5) 
```


**(3.) Bar plot of Raw Water for comparison at different locations**

```{r fig.height=10, fig.width=10}
psq_family_shared_RAW <- prune_taxa(shared_families, psq_family_RAW)
psq_family_shared_RAW <- prune_taxa(taxa_sums(psq_family_shared_RAW) > 0, psq_family_shared_RAW)
psq_family_shared_RAW

psq_family_shared_RAW_rel <- transform_sample_counts(psq_family_shared_RAW, function(x) x / sum(x))


RAW_bar <- plotbar(psq_family_shared_RAW_rel, group = "anon_location", level = "Phylum", top = 20) + 
  labs(y = "Relative Abundance", x = "Water Treatment Works", title = "Raw Water") +
  theme(axis.text.x = element_text(size = 12))

filename <- paste0(output_directory,"bar_plot_DCWW_Hurleston_RAW.svg")
ggsave(filename, RAW_bar, height = 10, width = 7.5)
```

```{r}
RAW_beta <- plotbeta(psq_family_shared_RAW_rel, group="anon_location") + 
  stat_ellipse(level = 0.95) +
  labs(colour = "Water Works") +
  scale_colour_manual(values = location_palette)

filename <- paste0(output_directory,"beta_diversity_DCWW_Hurleston_RAW.svg")
ggsave(filename, RAW_beta, height = 5, width = 7.5)
```

################################################################################
#                                                                              #
#                           **TAXONOMIC DOT PLOTS**                            #
#                                                                              #
################################################################################

**(1.) Overview of Hurleston**

```{r, fig.height=10, fig.width=20}
psq_RGF <-   subset_samples(psq_obj_Hurleston_Clean, Sample.details %in% c("RGF", "Backwash-Water", "Backwash"))
psq_RGF <- tax_glom(psq_RGF, taxrank = "Family")
psq_RGF <- prune_taxa(taxa_sums(psq_RGF) > 0, psq_RGF)
psq_RGF_rel <- transform_sample_counts(psq_RGF, function(x) x / sum(x))


psq_RGF_rel_filt <- prune_taxa(
  apply(otu_table(psq_RGF_rel), 1, max) >= 0.01,
  psq_RGF_rel
)


df <- psmelt(psq_RGF_rel_filt)

top20_families <- df %>%
  group_by(Family) %>%
  summarise(total_abundance = sum(Abundance, na.rm = TRUE)) %>%
  arrange(desc(total_abundance)) %>%
  slice_head(n = 20) %>%
  pull(Family)

df_top20 <- df %>% filter(Family %in% top20_families)

df_top20$Phylum <- factor(df_top20$Phylum,
                         levels = unique(df_top20$Phylum))

df_top20$Family <- factor(df_top20$Family, levels = sort(top20_families))


phylum_blocks <- df %>%
  group_by(Phylum) %>%
  summarise(
    y_min = min(as.numeric(Family)),
    y_max = max(as.numeric(Family))
  )

df_top20$Sample <- factor(df_top20$Sample)

df_top20

df_top20$batch.name <- recode(
  df_top20$batch.name,
    "RGF1B_1L_Matrix" = "Backwash Matrix",
    "RGF1B_Matrix" = "Backwash Matrix", 
  "RGF1B_Water" = "Backwash Water", 
    "RGF1" = "RGF1",
    "RGF2" = "RGF2",
    "RGF3" = "RGF3",
    "RGF4" = "RGF4",
    "RGF5" = "RGF5",
    "RGF6" = "RGF6",
    "RGF7" = "RGF7"
  )

  Dot_Hurleston <- ggplot(
  df_top20,
  aes(
    x = Sample,
    y = Family,
    size = Abundance,
    color = Family
  )
) +
  geom_point(alpha = 0.8) +
  theme_bw() +
  xlab("") +
  ylab("Family") +

  ggh4x::facet_nested(
    rows = vars(Phylum, Family),
    cols = vars(Sample.details, batch.name),
    scales = "free",
    space = "free",
    switch = "both"
  ) +

  scale_size_continuous(range = c(0, 12)) +

  scale_x_discrete(
    labels = setNames(df_top20$batch.name, df_top20$Sample)
  ) +

  guides(
    color = "none",
    size  = guide_legend(
      title = "Relative\nAbundance",
      override.aes = list(color = "grey30")
    )
  ) +

  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    strip.text = element_text(size = 10, face = "bold", lineheight = 0.9),
    strip.text.y.left = element_text(angle = 0),
    strip.background = element_rect(
      colour = "grey30",
      fill = alpha("cornflowerblue", 0.5)
    ),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "right",
    plot.title.position = "plot"
  ) +
  ggtitle("         Phylum                   Family")


filename <- paste0(output_directory,"dot_plot_Hurleston.svg")
ggsave(filename, Dot_Hurleston, height = 15, width = 15)

```

################################################################################
#                                                                              #
#            **RANDOM FOREST FOR PREDICTION T&O BIOLOGICAL REMOVAL**           #
#                                                                              #
################################################################################

**(1.) Comparative table of shared communities at each taxonomic level**

```{r}

psq_obj_ref_rel <- transform_sample_counts(psq_obj_ref, function(x) x / sum(x))

# Define the taxonomic ranks you want
tax_levels <- c("Phylum", "Class", "Order", "Family", "Genus", "ASV")

# Empty list to save results
results <- list()

for (rank in tax_levels) {

  # ---- (1) Agglomerate or keep ASV level ----
  if (rank == "ASV") {
    ps_rank <- psq_obj_ref_rel
  } else {
    ps_rank <- tax_glom(psq_obj_ref_rel, taxrank = rank)
  }
  
  # ---- (1.5) Subset for total amount ----
  taxa_all <- taxa_names(ps_rank)
  
  # ---- (2) Subset Hurleston ----
  ps_hurleston <- subset_samples(ps_rank, Location == "Hurleston")
  ps_hurleston <- prune_taxa(taxa_sums(ps_hurleston) > 0, ps_hurleston)
  taxa_h <- taxa_names(ps_hurleston)
  
  # ---- (3) Subset DCWW (non-Hurleston) ----
  ps_dcww <- subset_samples(ps_rank, Location != "Hurleston")
  ps_dcww <- prune_taxa(taxa_sums(ps_dcww) > 0, ps_dcww)
  taxa_d <- taxa_names(ps_dcww)
  
  # ---- (4) Calculate unique + shared ----
  unique_h <- setdiff(taxa_h, taxa_d)
  unique_d <- setdiff(taxa_d, taxa_h)
  shared   <- intersect(taxa_h, taxa_d)
  
  # ---- (5) Save results ----
  results[[rank]] <- data.frame(
    Level = rank,
    Total = length(taxa_all),
    Hurleston = length(unique_h),
    DCWW = length(unique_d),
    Shared = length(shared)
  )
}

# Bind into a single dataframe
final_results <- bind_rows(results)
final_results
```

```{r}
# (1.) IDENTIFYING TAXA SHARED BETWEEN THE DCWW AND HURLESTON AT THE FAMILY LEVEL
psq_family <- tax_glom(psq_obj_ref, taxrank = "Family")      # Reduce object to Family Level

# (3.) EXTRACT DATA WHICH ISNT HURLESTON
ps_dcww <- subset_samples(psq_family, !Location %in% c("Hurleston", "Cefni", "Glascoed", "Preseli"))
ps_dcww <- subset_samples(ps_dcww, batch.name %in% c("oct_19", "dec_19"))
ps_dcww <- subset_samples(ps_dcww, Sample.details == "RGF")
ps_dcww <- prune_taxa(taxa_sums(ps_dcww) > 0, ps_dcww)

```

```{r}
## Extract OTU Table from Phyloseq Object
reference_bact_otu <- data.frame(t(otu_table(ps_dcww)))

## Extract Metadata from Phyloseq Object
metadata_ref       <- data.frame(sample_data(ps_dcww))
```

**NB: These optimisation plots may take some time to run depending on the size of the otu dataframe.**

(1.)
```{r}
plot(randomForest(reference_bact_otu, y=metadata_ref$RGF.TO.removal, proximity = F, keep.forest = F, ntree=100000))
```
(2.)
* Replace ntree with an optimal value from the previous plot. 
```{r}
cv.treat<-rfcv(reference_bact_otu, metadata_ref$RGF.TO.removal, cv.fold = 100, ntree = 5000)
plot(cv.treat$error.cv~cv.treat$n.var, type = "S", xlab = "No. variables", ylab = "Error rate", ylim = c(0,1), main = "RFCV")
```

### Running the RF with optimal settings. 

```{r}

set.seed(2006576)
#Run a random forest classification for surface
TO_RF<-randomForest(reference_bact_otu, y=metadata_ref$RGF.TO.removal, importance = T, proximity = T, mtry = 30, ntree=5000)

#Look at how accurately the classification worked
TO_RF$confusion
```

```{r, fig.height=5, fig.width=5}
set.seed(08121001)

## Assigning Colours
# Ellipses
ellipse_cols <- c(
  "NO"  = "brown3",
  "YES" = "cornflowerblue",
  "PREDICTION" = "green4"
)

ellipse_cols <- ellipse_cols[levels(metadata_ref$RGF.TO.removal)]

# Points as Location
# Factor of locations
loc_factor <- as.factor(metadata_ref$Location)
# Colour palette (distinct colours for each location)
location_cols <- brewer.pal(n = length(levels(loc_factor)), name = "Dark2")
# Create named palette
names(location_cols) <- levels(loc_factor)

## Shape as Media
shape_values <- c(
  "Anthracite" = 17,
  "GAC" = 15,
  "Sand" = 16
)



shape_factor <- as.factor(metadata_ref$RGF.media)
names(shape_values) <- levels(shape_factor)


#Do an ordination plot based on the random forest
RFplot.surface<-MDSplot(TO_RF, metadata_ref$RGF.TO.removal, xlim=c(-1,1), ylim=c(-1,1))

#svg("C:/Users/Danie/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/Add to Github/16S-Microbiome-Analysis/random_forest_MDS.svg")
plot(
  RFplot.surface$points,
  col = location_cols[loc_factor],     # <-- COLOUR BY LOCATION
  pch = shape_values[shape_factor],
  xlab = "MDS1",
  ylab = "MDS2",
  xlim = c(-1, 1),
  ylim = c(-1, 1)
)


#Add 95% confidence intervals
col = ellipse_cols[metadata_ref$RGF.TO.removal]

ordiellipse(RFplot.surface, metadata_ref$RGF.TO.removal,  col = ellipse_cols, conf=0.95,lwd=3,lty=3,draw="lines")

#Add a legend
# Ellipse Legend
legend("bottomleft",  c(levels(metadata_ref$RGF.TO.removal)),  col = ellipse_cols, pch=c(20), bty="n", cex=1)

# Location Legend
legend("bottomright", legend = names(location_cols), col = location_cols, pch = 19, bty = "n", cex = 1)

# Shape Legend
legend("topright", legend = names(shape_values), pch = shape_values, bty = "n", cex = 1)

#dev.off()
```

