---
title: "16S Bacterial Community Analysis 2026 - IoM"
author: "Dan Shelley"
date: "`r Sys.Date()`"
output: html_document
---

################################################################################
#                                                                              #
#            **IT IS RECOMMENDED TO RUN THIS RMD CHUNK BY CHUNK**              #
#                                                                              #
#         **THIS MARKDOWN CAN ALSO FUNCTION AS A CLASSIC 16S ANALYSIS**        #
#                                                                              #
################################################################################

################################################################################
#                                                                              #
#                       **Methodology & Package Summary**                      #
#                                                                              #
#        * Phyloseq is used to generate a Phyloseq object from Qiime2          #
#          outputs.                                                            #
#                                                                              #
#        * Taxa filtering: No Filtering for Taxa                               #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
#                                                                              #
################################################################################

################################################################################
#                                                                              #
#                            **Loading Libraries**                             #
#                                                                              #
################################################################################

```{r library.loading, include=FALSE}
#devtools::install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
devtools::install_github("gauravsk/ranacapa")

library(tidyverse)          #v1.2.0   - Used for data cleaning and manipulation.
library(vegan)          #v2.7-3   - Plotting Random Forest MDS.
library(ggplot2)        #v4.0.2   - Generic Plotting.
library(ggpubr)         #v        - 
library(FSA)            #v        - 
library(stringr)        #v        - 
library(ggh4x)          #v        -
library(phyloseq)       #v1.50.0  - Generating objects to store Qiime2 outputs.
library(microbial)      #v        - Phyloseq disperision statistical test.
library(pairwiseAdonis) #v        - Pairwise PERMANOVA functions.
```

################################################################################
#                                                                              #
#                          **Defining Input Files**                            #
#                                                                              #
################################################################################

```{r}
ref_metadata_path        <- "C:/Users/Danie/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/IoM_Nov_2025/IoM_metadata.csv"
ref_otu_table_path       <- "C:/Users/Danie/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/IoM_Nov_2025/qiime2output/exported_data/feature-table.tsv"
ref_taxonomy_table_path  <- "C:/Users/Danie/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/IoM_Nov_2025/qiime2output/exported_data/taxonomy.tsv"
ref_tree_path            <- "C:/Users/Danie/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/IoM_Nov_2025/qiime2output/exported_data/tree.nwk"

output_directory <- "C:/Users/Danie/OneDrive - Cardiff University/004 - Cardiff University - Senior Research Technician/IoM_Nov_2025/16S_Output/"
```

################################################################################
#                                                                              #
#                       **Generating Phyloseq Object**                         #
#                                                                              #
################################################################################

**NB: Remove the first line #BIOM and #OTU ID from the Feature Table Files prior to running**

```{r read_in_reference_data}
feature_table <- read.table(
  ref_otu_table_path,
  header = TRUE,
  row.names = 1,
  sep = "\t"
)

feature_table <- as.matrix(feature_table)
feature_table <- otu_table(feature_table, taxa_are_rows = TRUE)


taxonomy_table <- read.table(
  ref_taxonomy_table_path,
  sep = "\t",
  header = TRUE,
  row.names = 1
)

taxonomy_split <- strsplit(as.character(taxonomy_table$Taxon), ";")

taxonomy_split_formatted <- lapply(taxonomy_split, function(x) {
  length(x) <- 7
  return(x)
})

taxonomy_mtx <- as.matrix(do.call(rbind, taxonomy_split_formatted))

# Remove prefixes like "d__", "p__" etc.
taxonomy_mtx <- sub("^[a-z]__*", "", taxonomy_mtx)  # removes d__, p__, etc.
taxonomy_mtx[taxonomy_mtx == ""] <- NA # now we fill in the remaining empty gaps with NAs so R doesn't get confused

# Set column names (use "Domain" or "Kingdom" as you prefer)
colnames(taxonomy_mtx) <- c("Domain",
                            "Phylum",
                            "Class",
                            "Order",
                            "Family",
                            "Genus",
                            "Species")
# Add back the ASV IDs which were lost from our original data frame
rownames(taxonomy_mtx) <- rownames(taxonomy_table)

# Convert to phyloseq tax_table object
taxonomy_table <- phyloseq::tax_table(taxonomy_mtx)

metadata_table <- read.csv(
  ref_metadata_path,
  row.names = 1,
  na.strings = "",
  stringsAsFactors = TRUE
)

# Match Metadata Sample IDs to those in the feature table
metadata_table <- metadata_table[rownames(metadata_table) %in% colnames(feature_table), ]
metadata_table$Sample.name <- rownames(metadata_table)

metadata_table <- metadata_table %>%
  mutate(
    type = as.character(type),
    type = if_else(
      type == "shell_biofilm" & bioreactor %in% c("S10", "S3"),
      paste0(type, "_", bioreactor),
      type
    ),
    type = factor(type)
  )

metadata_table <- metadata_table %>%
  mutate(
    type = as.character(type),
    type = if_else(
      type == "outflow_drainage" & bioreactor %in% c("S10", "S3", "L"),
      paste0(type, "_", bioreactor),
      type
    ),
    type = factor(type)
  )

metadata_table$matrix_type <- metadata_table$matrix_type |>
  as.character() |>
  recode(
    "bulk" = "Bulk",
    "matrix" = "Matrix",
    "precipitate" = "Precipitate",
    .missing = "Other"
  )

metadata_table <- sample_data(metadata_table)

rooted_tree <- read_tree(ref_tree_path)

psq_obj_ref <- phyloseq(feature_table, taxonomy_table, metadata_table, rooted_tree)
```

################################################################################
#                                                                              #
#                              **Data Cleaning**                               #
#                                                                              #
################################################################################

```{r}
psq_obj_ref_no_small <- subset_samples(psq_obj_ref, !type %in% c("shell_biofilm_S3","outflow_drainage_S3"))

## Removing taxa where total counts are <10
physeq_filtered <- prune_taxa(
  taxa_sums(psq_obj_ref_no_small) >= 10,
  psq_obj_ref_no_small
)

sample_data(physeq_filtered)$Sequencing.depth <- sample_sums(physeq_filtered)
```

```{r check_ref_psq}

sample_data(physeq_filtered)$type <- recode(
  sample_data(physeq_filtered)$type,
  "inflow_water" = "Influent",
  "outflow_water" = "Effluent",
  "outflow_drainage_S10" = "ED Sludge Shell",
  "outflow_drainage_L" = "ED Sludge Limestone",
  "shell_biofilm_S10" = "Shell",
  "limestone_biofilm" = "Limestone",
  "header_water" = "Header Tank"
)

sample_data(physeq_filtered)$type <- as.character(sample_data(physeq_filtered)$type)
sample_data(physeq_filtered)$type <- factor(
  sample_data(physeq_filtered)$type,
  levels = c(
    "Influent",
    "Header Tank",
    "Effluent",
    "Shell",
    "ED Sludge Shell",
    "Limestone",
    "ED Sludge Limestone"
  )
)

sample_data(physeq_filtered)$types <- recode(
  sample_data(physeq_filtered)$types,
  "inflow_water" = "Inflow_Water",
  "outflow_drainage" = "Outflow_Drainage",
  "shell_biofilm_bulk" = "Bulk",
  "shell_biofilm_matrix" = "Matrix",
  "limestone_biofilm_bulk" = "Bulk",
  "limestone_biofilm_matrix" = "Matrix",
  "header_water" = "Header_Water",
  "limestone_biofilm_ppt" = "Precipitate",
  "outflow_water" = "Outflow_Water",
  "shell_biofilm_ppt" = "Precipitate"
)

## Removing Control Groups
physeq_filtered <- subset_samples(physeq_filtered, control != "Y")
```

```{r}

#hcl.colors(9,"Dark2", alpha = 1)
group_colours <- c(
  "Influent"              = "#C87A8AFF",  # deep blue
  "Header Tank"              = "#B9865FFF",  # neutral dark grey
  "Effluent"             = "#9A9346FF",  # muted orange

  "Shell"                       = "#00A1A9FF",  # soft green
  "ED Sludge Shell"    = "#5F96C2FF",  # deeper green

  "Limestone"                       = "#9E85C5FF",  # soft purple
  "ED Sludge Limestone"    = "#C179B0FF"   # deeper purple
)


types_shape <- c(
  "Bulk" = 15,
  "Matrix" = 17,
  "Precipitate" = 18,
  "Other" = 16
)
```



################################################################################
#                                                                              #
#       psq_obj_ref              : Full IoM Data, no small reactor             #
#                                                                              #
################################################################################


################################################################################


################################################################################
#                                                                              #
#                             **ALPHA DIVERSITY**                              #
#                                                                              #
################################################################################

**(1.) ALPHA RAREFRACTION CURVE**

```{r}

# Define depths
lib_sizes <- sample_sums(physeq_filtered)
min_lib   <- min(lib_sizes)
max_lib   <- max(lib_sizes)

# Start with your candidate depths
depths <- seq(1, max_lib, by = 1000)
sample_var <- "Sample.name"

# Initialize list
richness_list <- list()

for (d in depths) {
  set.seed(2006576)
  rare_dat <- rarefy_even_depth(physeq_filtered, sample.size = d, rngseed = 2006576, verbose = FALSE, replace = FALSE)

  richness <- estimate_richness(rare_dat, measures = "Observed")
  richness$SampleID <- rownames(richness)
  richness$Depth <- d
  richness_list[[as.character(d)]] <- richness
}

# Combine results
rare_df <- bind_rows(richness_list)

# Add metadata
metadata <- data.frame(sample_data(physeq_filtered))
metadata$SampleID <- rownames(metadata)
rare_df <- left_join(rare_df, metadata, by = "SampleID")

# Summarize mean & SE per group
summary_df <- rare_df %>%
  group_by(Depth, !!sym(sample_var)) %>%
  mutate(
    Richness_mean = mean(Observed, na.rm = TRUE),
    Richness_se   = sd(Observed, na.rm = TRUE) / sqrt(n())
  ) %>%
  ungroup()
```
**(2.) Plotting Alpha Rarefraction Curve**

```{r}

Rare.plot <- ggplot(summary_df, aes(x = Depth, y = Richness_mean, group = SampleID, color = type)) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = group_colours) +
  scale_x_continuous(breaks = seq(0, 100000, by = 5000)) +
  labs(x = "Sequencing Depth (n)", y = "Mean Observed ASVs (n)", color = "Stage") +
  theme_bw()

filename <- paste0(output_directory,"rarefraction_curve_Hurleston.svg")
ggsave(filename, Rare.plot, height = 5, width = 10)
```

```{r}
physeq_filtered <- rarefy_even_depth(physeq_filtered, 20000, rngseed = 202020)
```



**(3.) Calculating Alpha Diversity Distances**

```{r}

group      <- "type"   # column used for grouping
ps11       <- physeq_filtered       # your phyloseq object


# Extract sample metadata
mapping <- phyloseq::sample_data(ps11)

# Make sure all taxa have some non-zero counts
ps11 <- phyloseq::filter_taxa(ps11, function(x) sum(x) > 0, prune = TRUE)

# Rename the grouping column for clarity
colnames(mapping) <- gsub(group, "Comparative_Group", colnames(mapping))

# Create final grouping variable
mapping$Group <- as.factor(mapping$Comparative_Group)

# Convert OTU table into a standard count matrix (samples x taxa)
count <- as.data.frame(t(ranacapa::vegan_otu(ps11)))

# Transpose back for vegan functions that expect taxa as columns
x <- t(count)

# Shannon diversity
Shannon <- vegan::diversity(x, index = "shannon")

# Inverse Simpson
Inv_Simpson <- vegan::diversity(x, index = "invsimpson")

# Species richness (observed)
Richness <- vegan::specnumber(x)

# Prevalence (alternative richness measure)
S2 <- rowSums(x > 0)

# Pielou's evenness
Pielou_evenness <- Shannon / log(Richness)

# Simpson evenness
Simpson_evenness <- Inv_Simpson / Richness

# Chao1, ACE, and richness from estimateR
est <- vegan::estimateR(x)
ACE   <- est["S.ACE", ]

# Combine all metrics into a single table
report <- cbind(
  Shannon,
  Inv_Simpson,
  Pielou_evenness,
  Simpson_evenness,
  Richness,
  ACE
)

alp <- merge(mapping, report, by = "row.names", all = FALSE)

# Store complete alpha-diversity table
alpha_table <- alp

# Select metrics to test
metrics_to_test <- c(
  "Shannon", "Inv_Simpson", "Pielou_evenness",
  "Simpson_evenness", "Richness", "ACE"
)


data_for_test <- cbind(
  data.frame(
    ID = seq_len(nrow(alpha_table)),
    Group = alpha_table$Group,
    SampleID = alpha_table$Row.names,
    Sequencing.depth = alpha_table$Sequencing.depth
  ),
  alpha_table[, metrics_to_test]
)

```

## Linear Model for Distance vs Sequencing Depth

**(4.) Plotting Distance Matrices**

```{r fig.height=7, fig.width=7}
data_for_test

kruskal.test(Richness ~ Group, data = data_for_test)

sigdf <- pairwise.wilcox.test(
  x = data_for_test$Shannon,
  g = data_for_test$Group,
  p.adjust.method = "BH"
)


pw_df <- as.data.frame(as.table(sigdf$p.value)) %>%
  rename(
    group1 = Var1,
    group2 = Var2,
    p = Freq
  ) %>%
  filter(!is.na(p))


pw_df <- pw_df %>%
  mutate(
    p.signif = case_when(
      p <= 0.0001 ~ "****",
      p <= 0.001  ~ "***",
      p <= 0.01   ~ "**",
      p <= 0.05   ~ "*",
      TRUE        ~ "ns"
    )
  )


pw_sig <- pw_df %>%
  filter(p.signif != "ns")


y_max <- max(data_for_test$Shannon, na.rm = TRUE)
step  <- 0.05 * y_max

pw_sig <- pw_sig %>%
  arrange(p) %>%
  mutate(y.position = y_max + row_number() * step)


alpha_plot <- ggplot(data_for_test, aes(x = Group, y = Shannon, fill = Group)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  stat_boxplot(geom = "errorbar", width = 0.5) +
  geom_point(aes(color = Group), size = 3, alpha = 0.6) +
  scale_fill_manual(values = group_colours) +
  scale_color_manual(values = group_colours) +
  labs(fill = "", x = "", y = "Shannon Index") +
  ylim(0,8) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 35, vjust = 1, hjust=1)) +
  guides(color = "none") +
  ggpubr::stat_pvalue_manual(
    pw_sig,
    label = "p.signif",
    tip.length = 0.01
  )

filename <- paste0(output_directory,"Alpha_Shannon.svg")
ggsave(filename, alpha_plot, height = 5, width = 7.5)


```


################################################################################
#                                                                              #
#                              **BETA  DIVERSITY**                             #
#                                                                              #
################################################################################

**(1.) IoM Data Normalisation**

```{r}
psq_obj_ref_rel <- normalize(physeq_filtered, method = "relative")

psq_obj_ref_rel <- prune_taxa(taxa_sums(psq_obj_ref_rel) > 0, psq_obj_ref_rel)

IoM_beta <- plotbeta(psq_obj_ref_rel, group = "type", shape = "matrix_type") + 
  labs(colour = "",
       shape = "") +
  stat_ellipse(aes(group = type), lty = "dashed") +
  scale_shape_manual(values = types_shape) +
  scale_colour_manual(values = group_colours)

filename <- paste0(output_directory,"beta_diversity_Hurleston.svg")
ggsave(filename, IoM_beta, height = 5, width = 7.5)
```

**(2.) PERMANOVA: Statistically comparing microbial community similarities**

```{r}

ps_tr <- microbiome::transform(psq_obj_ref_rel, "clr")
ps_dist_matrix <- phyloseq::distance(psq_obj_ref_rel, method ="bray")

sample_data(ps_tr)$batch.name <- as.character(sample_data(ps_tr)$type)
ps_tr <- subset_samples(ps_tr, type !="NA")

vegan::adonis2(ps_dist_matrix ~ phyloseq::sample_data(ps_tr)$type)
pairwise.adonis(ps_dist_matrix, phyloseq::sample_data(ps_tr)$type)
permutest(betadisper(ps_dist_matrix, group = phyloseq::sample_data(ps_tr)$type), pairwise = TRUE)

```

################################################################################
#                                                                              #
#                           **TAXONOMIC BAR PLOTS**                            #
#                                                                              #
################################################################################

**(1.) IoM**

```{r, fig.height=7.5, fig.width=7.5}

plotbar(psq_obj_ref_rel, level = "Family", group = "type", top = 20, fontsize.x = 7.5, fontsize.y = 7.5) +
  labs(y = "Relative Abundance", x = "Sample")

```



################################################################################
#                                                                              #
#                           **TAXONOMIC DOT PLOTS**                            #
#                                                                              #
################################################################################

**(1.) Overview of Hurleston**

```{r, fig.height=10, fig.width=7.5}
physeq_filtered_gen <- tax_glom(physeq_filtered, taxrank = "Genus")
psq_obj_ref_rel <- normalize(physeq_filtered_gen, method = "relative")
psq_obj_ref_rel <- prune_taxa(taxa_sums(psq_obj_ref_rel) > 0, psq_obj_ref_rel)


df_psq_fam <- psmelt(psq_obj_ref_rel)

genera <- c(
  "Terrimonas",
  "Stenotrophomonas",
  "Sphingorhabdus",
  "Sphingopyxis",
  "Sphingomonas",
  "Shewanella",
  "Reyranella",
  "Pseudonocardia",
  "Pseudomonas",
  "Pedobacter",
  "Novosphingobium",
  "Nocardia",
  "Nitrospira",
  "Mycobacterium",
  "Microbacterium",
  "Methylobacterium",
  "Leptolyngbya",
  "Hyphomicrobium",
  "Flavobacterium",
  "Duganella",
  "Caulobacter",
  "Brevundimonas",
  "Brevibacterium",
  "Bosea",
  "Afipia"
)


df_psq_fam_filtered <- df_psq_fam %>%
  filter(Genus %in% genera)


df_psq_fam_filtered <- df_psq_fam_filtered %>%
  filter(Abundance > 0)


mn_plot <-ggplot(
  df_psq_fam_filtered,
  aes(
    x = Sample,
    y = Genus,
    size = Abundance,
    color = Genus
  )
) +
  geom_point(alpha = 0.8) +
  facet_wrap(~ type, scales = "free_x") +
  theme_bw() +
  xlab("") +
  ylab("Genus") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
filename <- paste0(output_directory,"Mn_Tax_plot.svg")
ggsave(filename, mn_plot, height = 10, width = 7.5)

######## Potential Nitrifying 
df_psq_fam <- psmelt(psq_obj_ref_rel)

taxa_keep <- c(
  "Candidatus_Anammoximicrobium",
  "Nitrosarchaeum",
  "Nitrosomonadaceae",
  "Nitrosomonas",
  "Nitrosospira",
  "Planctomycetota"
)


psq_subset <- subset_taxa(
  psq_obj_ref_rel,
  Genus %in% taxa_keep |
  Family %in% taxa_keep |
  Phylum %in% taxa_keep
)

df_limited <- psmelt(psq_subset)


df_limited <- df_limited %>%
  mutate(
    taxa = case_when(
      !is.na(Genus)  ~ Genus,
      !is.na(Family) ~ Family,
      !is.na(Phylum) ~ Phylum,
      TRUE           ~ NA_character_
    )
  )

df_limited <- df_limited %>%
  filter(Abundance > 0)

Nit_plot <-ggplot(
  df_limited,
  aes(
    x = Sample,
    y = Family,
    size = Abundance,
    color = Family
  )
) +
  geom_point(alpha = 1) +
  facet_wrap(~ type, scales = "free_x") +
  theme_bw() +
  xlab("") +
  ylab("Genus") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

############# Micritisation


micritization_genera <- c(
  "Mastigocoleus",
  "Tessarococcus",
  "Halomonas",
  "Virgibacillus",
  "Desulfovibrio",
  "Desulfosarcina",
  "Desulfococcus",
  "Sporosarcina",
  "Achromatium",
  "Methylosinus"
)

df_psq_fam_filtered <- df_psq_fam %>%
  filter(Genus %in% micritization_genera)


df_psq_fam_filtered <- df_psq_fam_filtered %>%
  filter(Abundance > 0)


mit_plot <-ggplot(
  df_psq_fam_filtered,
  aes(
    x = Sample,
    y = Genus,
    size = Abundance,
    color = Genus
  )
) +
  geom_point(alpha = 0.8) +
  facet_wrap(~ type, scales = "free_x") +
  theme_bw() +
  xlab("") +
  ylab("Genus") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
